---
title: "Hypothesis 1"
author: "Zonghong Yu"

output: rmdformats::robobook
---



```{r}
# Packages
library(tidyverse)
library(plotly)
library(ggplot2)
library(boot)
```

Write your Null and Alternative Hypothesis.
```{r}
df <- read.csv("spotify_cleaned.csv",stringsAsFactors=FALSE)
head(df)

# My Data Science Question: 
# Is the population mean of popularity of rock music greater than that of rap music?


# Hypothesis:
# H0: The mean popularity of rock music is the same as the mean popularity of rap music
# Ha: The mean popularity of rock music is higher than the mean popularity of rap music
```

Do a t-test (if your Data science question is about population averages. If your question is about comparing proportions then use a Z-test), and write your conclusion at 5% significance level.
```{r}
# Since my question is to compare average, therefore, I use t-test for my hypothesis:

dt_pop <- df$popularity[df$genre == "rock"]


ur_pop <- df$popularity[df$genre == "rap"]




# Since the hypothesis is to compare energy of KC and AG,  alternative is higher, use greater in t-test.
t.test(ur_pop, dt_pop, alt="greater",conf.level = 0.95)  #at 5% significance level.
```
```{r}
# Conclusion:
# Based on the t-test, we can see that the p-value is less than 0.05 which is the 5% significance level.
# Therefore, we reject the Null hypothesis at this significant level.
# We can conclude that we have enough evidence to prove that the mean energy of Kelly Clarkson is not the same as the mean energy of Ariana Grande. In fact, we can say that the mean energy of Kelly Clarkson is larger than the mean energy of Ariana Grande

# this conclusion also aligns with my EDA analysis, that the mean energy of Kelly Clarkson is larger than the mean energy of Ariana Grande
```


Do a bootstrap test(here you will be using bootstrap sampling and 95% bootstrap percentile interval) to answer the same question and write your conclusion at 5% significance level.

```{r}

# Use bootstrap

set.seed(2)
difference <- rep(NA,10000)
boot_ratio <- rep(NA, 10000)

for (j in 1:10000){
  boot_dt <- mean(sample(dt_pop, length(dt_pop), replace = T))
  boot_ur <- mean(sample(ur_pop, length(ur_pop), replace = T))
  difference[j] <- boot_ur - boot_dt   #the difference
  boot_ratio[j] <- boot_ur / boot_dt # The ratio
}

mean(difference) #bootstrap mean difference
mean(boot_ratio) #bootstrap mean rAtio
```
```{r}
hist(difference, main = "Bootstrap distribution of difference in means for popularities of two genres",col = '#FFFF99')
abline(v = mean(ur_pop) - mean(dt_pop), col = "blue", lty = 2)
```
```{r}
qqnorm(difference)
qqline(difference)
```
```{r}
# 95% CI
CI <- quantile(difference, c(0.025, 0.975))
CI
```
```{r}
hist(difference, main = "Bootstrap of difference in means for popularities of two genres",col = '#FFFF99')
abline(v = mean(ur_pop) - mean(dt_pop), col = "blue", lty = 2)
abline(v = CI, col = 4, lwd = 2)
```







